Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-11-15 paper/paper.tex
merged validation flow

\centerline{\includegraphics[width=0.85\textwidth]{figures/Validation_Figure/ifn_plot.pdf}}
\caption{Validation that IFN induction is increased by some of the mutations identified in the single-cell virus sequencing of IFN+ cells.
{\bf (A)}
Percent of infected cells that become IFN+ after infection with a bulk stock of the indicated viral mutant, as determined using a reporter cell line (see \SUPPFIG{SNPflow} for details).
The numbers indicate the median of four measurements for each viral mutant.
The limit of detection of 0.05\% is indicated with a dashed green line, and the median value for wild type is indicated with a dashed blue line.
Points are colored orange if the mutant virus stock induces IFN more frequently than the wild-type viral stock (one-sided t-test, $P < 0.01$), and blue otherwise.
{\bf (B)}
Similar to the first panel, but validates the increased IFN induction for a large internal deletion in the PB1 gene.
See \SUPPFIG{delflow} and \SUPPFIG{delqPCR} for more detailed experimental data.
The experiments in the two panels were performed on different days, and so numerical values can be reliably compared within panels but not between panels.
}
\label{fig:validation}

@@

We generated the wild-type and point-mutant viruses for the validation experiments in \FIG{validation}A by reverse genetics using the pHW18* series of WSN reverse genetics plasmids~\citep{hoffmann2000dna}, but substituting the appropriate mutant plasmid listed in \FIGDATA[validation]{sequences} for the wild-type plasmid for that gene.
To generate the viruses from these plasmids, we transfected an equimolar mix of all eight plasmids into co-cultures of 293T and MDCK-SIAT1 cells seeded at a ratio of 8:1.
At 24 hours post-transfection, we changed media from D10 to influenza growth media. 
At 50 hours post-transfection (for the replicate 1 viruses in \SUPPFIG{SNPflow}) or 72 hours (for the replicate 2 viruses in \SUPPFIG{SNPflow}), we harvested the virus-containing supernatant, clarified this supernatant by centrifugation at 300$\times$g for 4 min, and stored aliquots of the clarified viral supernatant at -80$^{\circ}$C.
We then thawed aliquots and titered by TCID50 on MDCK-SIAT1 cells.
For the infections in \SUPPFIG{SNPflow}, we wanted to use equivalent particle counts, so we normalized all viruses to an equivalent hemagglutination titer on turkey red blood cells~\citep{hirst1942quantitative}.
 Briefly, a solution of 10\% v/v red blood cells (LAMPIRE Biological Laboratories, Fisher Scientific catalogue number 50412942)  was washed in PBS and diluted to a final concentration of 0.5\% v/v. 
Two-fold serial dilutions of virus were added to an equal volume of diluted red blood cells, and titer was measured as the highest dilution of viral stock at which complete hemagglutination of red blood cells was observed.
We then performed infections of the A549 reporter cell line at equivalent hemagglutination titer and analyzed the data as described in \SUPPFIG{SNPflow}.

To generate the NS1stop mutant virus and the wild-type and PB1del385to2163 mutant viruses in \SUPPFIG{delflow}, we used slightly different procedures.
The wild-type virus was generated by reverse genetics as described for the point-mutant viruses above, harvested at 48 hours post-transfection, and then passaged on MDCK-SIAT1 cells for 36 hours at an MOI of 0.05---conditions that we previously validated to lead to relatively little accumulation of defective particles~\citep{russell2018extreme}.
The NS1stop virus was similarly generated, but was passaged for 48 rather than 36 hours, since it had slower growth kinetics and so needed a longer period of time to reach high titers.
The viruses with deletions in the PB1 segment could not be generated in normal 293T and MDCK-SIAT1 cells, since they required the exogenous expression of the PB1 protein.
Therefore, these viruses were generated in previously described 293T and MDCK-SIAT1 cells that had been engineered to constitutively express PB1~\citep{bloom2010permissive}.
These viruses were harvested from transfections at 72 hours, and passaged twice in the MDCK-SIAT1 cells constitutively expressing PB1 at a MOI of 0.001 for 72 hours and 0.01 for 48 hours. 
This passaging was necessary as viral titers from transfections were too low to generate sufficient virus from a single passage.
The wild-type and NS1stop viruses were titered by TCID50 on MDCK-SIAT1 cells, and the PB1 deletion viruses were titered on the MDCK-SIAT1 cells constitutively expressing PB1.
The infections in \SUPPFIG{delflow} were performed at equivalent TCID50s as described in the legend to that figure.
That these equivalent TCID50s were also roughly equivalent in terms of particles capable of undergoing primary transcription is shown in \SUPPFIG{delqPCR}.

@@

\begin{suppfig}
\includegraphics[height=0.71\textheight]{figures/Validation_Figure/SNP_flow_plot.pdf}
\caption{Flow cytometry data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were not infected (three replicates) or infected with stocks of the indicated viral mutant in quadruplicate.
After 13 hours, cells were stained for expression of HA protein and analyzed by FACS for HA and the ZsGreen driven by the \textit{IFNL1} reporter.
The contour plots show the density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of the uninfected cells in these populations.
For all influenza-infected cells, the percentage IFN+ was calculated only among the HA+ cells (since these are the ones that are infected).
For the uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independently generated stocks were each assayed in duplicate (i.e., \#1a and \#1b are replicates of one viral stock, and \#2a and \#2b are replicates of the other viral stock).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
The HA staining indicates that the number of transcriptionally active viral units per cell was similar across most mutants, and calculating percentage IFN+ only among HA+ cells largely corrects for modest variation in titer across mutants.
}
\label{suppfig:SNPflow}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.7\textwidth]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{
Flow cytometry data for \FIG{validation}B.
Each infection was performed in triplicate.
A complicating factor for these data is that the virus with the deletion in PB1 (PB1del385to2163) cannot perform secondary transcription because it does not express PB1 protein, and so cannot be effectively normalized by HA expression since it expresses lower levels of HA due to the lack of secondary transcription.
Therefore, for the data in this figure all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for the wild-type and NS1stop viruses, and on MDCK-SIAT1 cells expressing PB1~\citep{bloom2010permissive} for the PB1del385to2163 virus.
Finally, the approach in \SUPPFIG{delqPCR} was used to show that equivalent TCID50s, all viral variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
All this normalization suggests that all infections were performed at a similar dose of particles active for primary transcription.
Therefore, the percent IFN+ was calculated from the flow data in this figure for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for the PB1del385to2163 virus.
}
\label{suppfig:delflow}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}
\caption{
Validation that the infections in \FIG{validation}B and \SUPPFIG{delflow} were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \SUPPFIG{delflow}), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins~\citep{killip2014activation}.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
\label{suppfig:delqPCR}
\end{suppfig}

>>>>

\centerline{\includegraphics[width=0.85\textwidth]{figures/Validation_Figure/ifn_plot.pdf}}
\caption{Validation that IFN induction is increased by some of the mutations identified in the single-cell virus sequencing of IFN+ cells.
{\bf (A)}
Percent of infected cells that become IFN+ after infection with a bulk stock of the indicated viral mutant, as determined using a reporter cell line.
The numbers indicate the median of four measurements for each viral mutant.
The limit of detection of 0.05\% is indicated with a dashed green line, and the median value for wild type is indicated with a dashed blue line.
Points are colored orange if the mutant virus stock induces IFN more frequently than the wild-type viral stock (one-sided t-test, $P < 0.01$), and blue otherwise.
{\bf (B)}
Similar to the first panel, but validates the increased IFN induction for a large internal deletion in the PB1 gene.
See \SUPPFIG{validationflow} and \SUPPFIG{delqPCR} for more detailed experimental data.
The experiments in the two panels were performed on different days, and so numerical values can be reliably compared within panels but not between panels.
}
\label{fig:validation}

@@

We generated the wild-type and point-mutant viruses for the validation experiments in \FIG{validation}A by reverse genetics using the pHW18* series of WSN reverse genetics plasmids~\citep{hoffmann2000dna}, but substituting the appropriate mutant plasmid listed in \FIGDATA[validation]{sequences} for the wild-type plasmid for that gene.
To generate the viruses from these plasmids, we transfected an equimolar mix of all eight plasmids into co-cultures of 293T and MDCK-SIAT1 cells seeded at a ratio of 8:1.
At 24 hours post-transfection, we changed media from D10 to influenza growth media. 
At 50 hours post-transfection (for the replicate 1 viruses in \SUPPFIG{validationflow}A) or 72 hours (for the replicate 2 viruses in \SUPPFIG{validationflow}), we harvested the virus-containing supernatant, clarified this supernatant by centrifugation at 300$\times$g for 4 min, and stored aliquots of the clarified viral supernatant at -80$^{\circ}$C.
We then thawed aliquots and titered by TCID50 on MDCK-SIAT1 cells.
For the infections in \SUPPFIG{validationflow}A, we wanted to use equivalent particle counts, so we normalized all viruses to an equivalent hemagglutination titer on turkey red blood cells~\citep{hirst1942quantitative}.
 Briefly, a solution of 10\% v/v red blood cells (LAMPIRE Biological Laboratories, Fisher Scientific catalogue number 50412942)  was washed in PBS and diluted to a final concentration of 0.5\% v/v. 
Two-fold serial dilutions of virus were added to an equal volume of diluted red blood cells, and titer was measured as the highest dilution of viral stock at which complete hemagglutination of red blood cells was observed.
We then performed infections of the A549 reporter cell line at equivalent hemagglutination titer and analyzed the data as described in \SUPPFIG{validationflow}A.

To generate the NS1stop mutant virus and the wild-type and PB1del385to2163 mutant viruses in \SUPPFIG{validationflow}B, we used slightly different procedures.
The wild-type virus was generated by reverse genetics as described for the point-mutant viruses above, harvested at 48 hours post-transfection, and then passaged on MDCK-SIAT1 cells for 36 hours at an MOI of 0.05---conditions that we previously validated to lead to relatively little accumulation of defective particles~\citep{russell2018extreme}.
The NS1stop virus was similarly generated, but was passaged for 48 rather than 36 hours, since it had slower growth kinetics and so needed a longer period of time to reach high titers.
The viruses with deletions in the PB1 segment could not be generated in normal 293T and MDCK-SIAT1 cells, since they required the exogenous expression of the PB1 protein.
Therefore, these viruses were generated in previously described 293T and MDCK-SIAT1 cells that had been engineered to constitutively express PB1~\citep{bloom2010permissive}.
These viruses were harvested from transfections at 72 hours, and passaged twice in the MDCK-SIAT1 cells constitutively expressing PB1 at a MOI of 0.001 for 72 hours and 0.01 for 48 hours. 
This passaging was necessary as viral titers from transfections were too low to generate sufficient virus from a single passage.
The wild-type and NS1stop viruses were titered by TCID50 on MDCK-SIAT1 cells, and the PB1 deletion viruses were titered on the MDCK-SIAT1 cells constitutively expressing PB1.
The infections in \SUPPFIG{validationflow}B were performed at equivalent TCID50s as described in the legend to that figure.
That these equivalent TCID50s were also roughly equivalent in terms of particles capable of undergoing primary transcription is shown in \SUPPFIG{delqPCR}.

@@

\begin{suppfig}
{\bf \Large A}\includegraphics[width=0.5\textwidth, valign=t]{figures/Validation_Figure/SNP_flow_plot.pdf}
\hspace{0.02\textwidth}
{\bf \Large B} \includegraphics[width=0.4\textwidth, valign=t]{figures/Validation_Figure/del_flow_plot.pdf}
\caption{Flow cytometry data for \FIG{validation}.
{\bf (A)}
Data for \FIG{validation}A.
A549 cells with the \textit{IFNL1} reporter driving LNGFR$\Delta$C-ZsGreen were infected with stocks of the indicated viral mutant.
After 13 hours, cells were stained for HA protein and analyzed by FACS.
Contour plots show density of all cells, and IFN+ cells are also indicated by orange dots.
Cells were classified as HA+ or IFN+ based on gates set to put 0.05\% of uninfected cells in these populations.
For influenza-infected cells, the percentage IFN+ was calculated only among the HA+ cells (since these are the ones that are infected).
For the uninfected cells, the percentage IFN+ was calculated among all cells, since uninfected cells do not express HA.
For each viral mutant, two independent stocks were assayed in duplicate (i.e., \#1a and \#1b are replicates of one viral stock, and \#2a and \#2b are replicates of the other).
The infections with replicate \#1 of the wild-type virus were performed at an MOI of 0.1 as determined by TCID50, and all other viruses were infected at an equivalent particle number as determined by HI assay. 
The HA staining indicates that the number of transcriptionally active viral units was similar across most mutants, and calculating percentage IFN+ only among HA+ cells corrects for modest variation in titer across mutants.
{\bf (B)}
Data for \FIG{validation}B.
The virus with the deletion in PB1 cannot be effectively normalized by HA expression since it expresses lower levels of HA due to the lack of secondary transcription.
Therefore, all cells were infected at an equivalent MOI of 0.3 as determined by TCID50 on MDCK-SIAT1 cells for the wild-type and NS1stop viruses, and on MDCK-SIAT1 cells expressing PB1~\citep{bloom2010permissive} for the PB1del385to2163 virus.
\SUPPFIG{delqPCR} shows that at these equivalent TCID50s, all variants had similar amounts of transcriptionally active virus in the absence of secondary transcription.
The percent IFN+ was calculated for \emph{all} cells (HA+ and HA-) since that is a more fair comparison for the PB1del385to2163 virus.
}
\label{suppfig:validationflow}
\end{suppfig}

\begin{suppfig}
\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}
\caption{
Validation that the infections in \FIG{validation}B and \SUPPFIG{validationflow}B were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \SUPPFIG{validationflow}B), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins~\citep{killip2014activation}.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
\label{suppfig:delqPCR}
\end{suppfig}
